\vspace{-13mm}\includegraphics[width=15mm,scale=3]{../images/IST_Logo.png} \\ {\fontsize{24}{16}Detection of Li in Technological Materials through Nuclear Reactions} \vspace{-5mm}

### 2024-01-19 text/paper/paper.tex
À
paper.tex

\paragraph{Lithium} Firstly, for the reaction \ref{eq:reactionLi}:

\begin{equation}
\begin{split}
  E_{\alpha_{0_1}} &= 7.6298 \text{ MeV} \\
  E_{\alpha_{0_2}} &= 11.0314 \text{ MeV} \\
  \theta_{\alpha_{0_1}} &= 165 ^{\circ} \\
  \theta_{\alpha_{0_2}} &= 12 ^{\circ} \\
  Q (\ce{^7Li}) &= 17.346 \text{ MeV} \\
\end{split}
\label{eq:energiesli}
\end{equation}

@@

\paragraph{Boron} Secondly, for the reaction \ref{eq:reactionB}:

\begin{equation}
\begin{split}
  E_{\alpha_0} &= 3.7372 \text{ MeV} \\
  E_{\ce{^8Be}} &= 3.1086 \text{ MeV} \\
  \theta_{\alpha_0} &= 165 ^{\circ} \\
  \theta_{\ce{^8Be}} &= 11 ^{\circ}\\
  Q (\ce{^11Br}) &= 8.591 \text{ MeV} \\
\end{split}
\label{eq:energiesBr}
\end{equation}

@@

We started by displaying the samples in a vertical sample holder with the given order: in the lowest position the implanted sample, followed by lithium aluminate ($\ce{LiAlO2}$) and finally lithium fluoride ($\ce{LiF}$). So that we can do a calibration from the number of counts per channel to energy, we put an additional source on the holder just above the lithium fluoride one. Using a ruler, we measured the distance between the samples so that afterwards we could control the lift in order to change the sample on the beam's target.

@@

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.35]{../../images/Calib.png}
    \captionof{figure}{Spectrum of the Calibration source}
\end{center}

@@

The equation used for the triple gaussian fit is:

\begin{equation}
    \begin{split}
    f(x) = a_0\exp{\left(-\frac{(x-\mu_0)^2}{\sigma_0}\right)} &+ a_1\exp{\left(-\frac{(x-\mu_1)^2}{\sigma_1}\right)} + \\
    &+ a_2\exp{\left(-\frac{(x-\mu_2)^2}{\sigma_2}\right)} + c
    \end{split}
    \label{eq:califit}
\end{equation}

We decided to do the analysis only with "detector 0". This anlaysis could be further corroborated using the "detector 1" data.
Using the triple gaussian, we obtained the following fits for "detector 0":

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.6]{../../images/TT_21_Chn0.png}
    \captionof{figure}{Channel 0 spectrum fit with [\ref{eq:califit}]}
\end{center}

@@

The fit parameters for means $\mu_0$, $\mu_1$ and $\mu_2$ in equation [\ref{eq:califit}] represent the energy value of each peak in channels. The computed values for these parameters, as well as the theoretical values of the energy in MeV for the peaks, are organised in the following tables:

\begin{table}[H]
\centering
\begin{tabular}{|w{0.8cm}|w{3cm}|w{2cm}|}
\hline
Peak & Channel 0 [Chn] & Theoretical Energy [keV] \\ \hline
1\textsuperscript{st} & $ \mu_0 = 557.00 \pm 0.11 $ & 5156.59 \\ \hline
2\textsuperscript{nd} & $ \mu_1 = 593.14 \pm 0.10 $ & 5485.56 \\ \hline
3\textsuperscript{rd} & $ \mu_2 = 627.44 \pm 0.10 $ & 5804.82 \\ \hline
\end{tabular}
\caption{Calibration Values for Channel 0}
\label{tab:calibration0}
\end{table}

@@

Using the values in these three tables we now need to compute a calibration equation for each of the channels.
We plotted the theoretical values of the energy of the three peaks as a function of fit parameters:

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.35]{../../images/Chn0_calib_with_Backscattering.png}
    \captionof{figure}{Calibration "detector 0"}
\end{center}

@@

\begin{center}
    \label{alfa_peaks}
    \centering
    \includegraphics[scale = 0.35]{../../images/AlphaPeaks.png}
    \captionof{figure}{Theoretical energy values of the two alfphas}
\end{center}

@@

\begin{center}
    \centering
    \includegraphics[scale = 0.35]{../../images/OverlapLiF_NC.png}
    \captionof{figure}{Spectrum of the $\ce{LiF}$ sample}
\end{center}

@@

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.4]{../../images/FullAnalysisLiF.png}
    \captionof{figure}{Full Analysis of $\ce{LiF}$ Spectrum}
\end{center}

@@

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.35]{../../images/OverlapLiAlO2.png}
    \captionof{figure}{Spectrum of the $\ce{LiAlO_2}$ sample}
\end{center}

@@

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.4]{../../images/FullAnalysisLiAlO2.png}
    \captionof{figure}{Full Analysis of $\ce{LiAlO_2}$ Spectrum}
\end{center}

@@

@@

Finally, we used the calibration to obtain the implanted sample energy spectrum overlapped with the expected $\alpha$ energy:
\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.35]{../../images/OverlapImplanted.png}
    \captionof{figure}{Spectrum of the implanted sample}
\end{center}

@@

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.4]{../../images/FullAnalysisImplanted.png}
    \captionof{figure}{Full Analysis of the Implanted Sample Spectrum}
\end{center}

@@

\section{Quantitative Analysis}
According to INSERT CHU REFERENCE it is possible to calculate the concentration of the element in the sample using with the following equations:
\begin{equation}
    N\tau_i=\frac{H_i}{\sigma (E_i)\Omega Q}
\end{equation}
\begin{equation}
    N\Delta x=\frac{A}{\sigma (E_i)\Omega Q}
\end{equation}

Where $N\tau_i$ are $N\Delta x$ different values of the superfitial concentration of the element. For comparison, we can approximate $N\Delta x \approx N\tau_i \cdot$\#channels.\\
$H_i$ is the number of counts in a FAT MUDAR spectrum peak (used in the $\ce{LiAlO2}$ and $\ce{LiF}$ samples), $A$ is the area of a singular peak (used in the implanted sample), $\sigma (E_i) $ is the cross section of the reaction, $\Omega$ is the solid angle of the detector, $Q$ is the max charge acquisition divided by $e$.

With this we get:\\
APROXIMACOES:\\
implanted: $1\times10^{17}$ atoms/cm$^2$\\
$\ce{LiAlO2}$: $3\times10^{18}$ atoms/cm$^2$\\
etc...\\

\section{Conclusion}

@@

\section{}

>>>>

\paragraph{Lithium} Firstly, for the reaction \ref{eq:reactionLi}:
\begin{equation}
\begin{split}
  E_{\alpha_{0_1}} &= 7.6298 \text{ MeV} \\
  E_{\alpha_{0_2}} &= 11.0314 \text{ MeV} \\
  \theta_{\alpha_{0_1}} &= 165 ^{\circ} \\
  \theta_{\alpha_{0_2}} &= 12 ^{\circ} \\
  Q (\ce{^7Li}) &= 17.346 \text{ MeV} \\
\end{split}
\label{eq:energiesli}
\end{equation}

@@

\paragraph{Boron} Secondly, for the reaction \ref{eq:reactionB}:
\begin{equation}
\begin{split}
  E_{\alpha_0} &= 3.7372 \text{ MeV} \\
  E_{\ce{^8Be}} &= 3.1086 \text{ MeV} \\
  \theta_{\alpha_0} &= 165 ^{\circ} \\
  \theta_{\ce{^8Be}} &= 11 ^{\circ}\\
  Q (\ce{^11Br}) &= 8.591 \text{ MeV} \\
\end{split}
\label{eq:energiesBr}
\end{equation}

@@

We started by displaying the samples in a vertical sample holder with the given order: in the lowest position the implanted sample, followed by lithium aluminate ($\ce{LiAlO2}$) and finally lithium fluoride ($\ce{LiF}$). In order to do a calibration from the number of counts per channel to energy, we added an additional source on the holder just above the lithium fluoride one. Using a pachymeter, we measured the distance between the samples so that afterwards we could control the hight in order to change the sample on the beam's target.

@@

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.3]{../../images/Calib.png}
    \captionof{figure}{Spectrum of the Calibration source}
\end{center}

@@

%The equation used for the triple gaussian fit is:
%\begin{equation}
%    \begin{split}
%    f(x) = a_0\exp{\left(-\frac{(x-\mu_0)^2}{\sigma_0}\right)} &+ a_1\exp{\left(-\frac{(x-\mu_1)^2}{\sigma_1}\right)} + \\
%    &+ a_2\exp{\left(-\frac{(x-\mu_2)^2}{\sigma_2}\right)} + c
%    \end{split}
%    \label{eq:califit}
%\end{equation}

We decided to do the analysis only with "detector 0". This analysis could be further corroborated using the "detector 1" data.
Using the triple gaussian, we obtained the following fit for "detector 0":

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.5]{images/TT_21_Chn0.png}
    \captionof{figure}{Detector 0 spectrum fit with triple gaussian}
\end{center}

@@

The fit parameters for the means, $\mu_0$, $\mu_1$ and $\mu_2$ obtained from the triple gaussian fit represent the energy value of each peak in channels. The computed values for these parameters, as well as the theoretical values of the energy, in MeV, for the peaks, are organised in the following tables:

\begin{table}[H]
\centering
\begin{tabular}{|w{0.8cm}|w{3cm}|w{2cm}|}
\hline
Peak & Detector 0 [Chn] & Theoretical Energy [keV] \\ \hline
1\textsuperscript{st} & $ \mu_0 = 557.00 \pm 0.11 $ & 5156.59 \\ \hline
2\textsuperscript{nd} & $ \mu_1 = 593.14 \pm 0.10 $ & 5485.56 \\ \hline
3\textsuperscript{rd} & $ \mu_2 = 627.44 \pm 0.10 $ & 5804.82 \\ \hline
\end{tabular}
\caption{Calibration Values for detector 0}
\label{tab:calibration0}
\end{table}

@@

Using the values in these three tables we now need to compute a calibration equation for each of the channels.
We plotted the theoretical values of the energy of the three peaks as a function of fit parameters:\vspace{-2mm}

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.3]{../../images/Chn0_calib_with_Backscattering.png}
    \captionof{figure}{Calibration "detector 0"}
\end{center}

@@

\begin{center}
    \label{alfa_peaks}
    \centering
    \includegraphics[scale = 0.3]{../../images/AlphaPeaks.png}
    \captionof{figure}{Theoretical energy values of the two alfphas}
\end{center}

@@

\begin{center}
    \centering
    \includegraphics[scale = 0.3]{../../images/OverlapLiF_NC.png}
    \captionof{figure}{Spectrum of the $\ce{LiF}$ sample}
\end{center}

@@

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.3]{../../images/FullAnalysisLiF.png}
    \captionof{figure}{Full Analysis of $\ce{LiF}$ Spectrum}
\end{center}

@@

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.3]{../../images/OverlapLiAlO2.png}
    \captionof{figure}{Spectrum of the $\ce{LiAlO_2}$ sample}
\end{center}

@@

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.3]{../../images/FullAnalysisLiAlO2.png}
    \captionof{figure}{Full Analysis of $\ce{LiAlO_2}$ Spectrum}
\end{center}

@@

(NÃO DEVIA ESTAR NA TEORIA?)

@@

Finally, we used the calibration to obtain the implanted sample energy spectrum overlapped with the expected $\alpha$ energy:
\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.3]{../../images/OverlapImplanted.png}
    \captionof{figure}{Spectrum of the implanted sample}
\end{center}

@@

\begin{center}
    \label{TT_21}
    \centering
    \includegraphics[scale = 0.3]{../../images/FullAnalysisImplanted.png}
    \captionof{figure}{Full Analysis of the Implanted Sample Spectrum}
\end{center}

@@

\section{Quantitative Analysis}
According to INSERT CHU REFERENCE it is possible to calculate the concentration of the element in the sample using with the following equations:\\
\begin{minipage}[t]{0.5\linewidth-0.5em} % some separation
    \vspace{0pt} % anchor for [t]
    \vspace{\dimexpr\ht\strutbox-\topskip}% remove excess glue
    \begin{equation}
        N\tau_i=\frac{H_i}{\sigma (E_i)\Omega Q}
    \end{equation}
\end{minipage}\hfill
\begin{minipage}[t]{0.5\linewidth-0.5em} % some separation
    \vspace{0pt} % anchor for [t]
    \begin{equation}
        N\Delta x=\frac{A}{\sigma (E_i)\Omega Q}
    \end{equation}
\end{minipage}

\vspace{4mm}

Where $N\tau_i$ are $N\Delta x$ different values of the superfitial concentration of the element. For comparison, we can approximate $N\Delta x \approx N\tau_i \times$\#channels.

$H_i$ is the hight, in number of counts, in a spectrum with accumulation (used in the $\ce{LiAlO2}$ and $\ce{LiF}$ samples), $A$ is the area of a singular peak (used in the implanted sample), $\sigma (E_i) $ is the cross section of the reaction, $\Omega$ is the solid angle of the detector, $Q$ is the max charge acquisition divided by $e$.

With this we get:\\
Implanted Sample: $N\Delta x = 9.25\times 10^{16}$ atoms per cm$^2$\\
$\ce{LiAlO2}$: $N\tau_i=4.31\times10^{17}$ atoms per cm$^2$\\
$\ce{LiF}$: $N\tau_i=1.15\times10^{18}$ atoms per cm$^2$

For comparison:
$\ce{LiAlO2}$: $N\Delta x=2.15\times10^{19}$ atoms per cm$^2$\\
$\ce{LiF}$: $N\Delta x=5.71\times10^{19}$ atoms per cm$^2$\\
From this, it is noticeable that the implanted sample has around 200x less Lithium than the $\ce{LiAlO2}$ and 600x less Lithium than the $\ce{LiF}$, as expected.

\section{Conclusion}
To conclude, we ended up with a successful experiment. We were able to detect the presence of lithium in all the 
analysed samples trough nuclear reactions. Furthermore, we managed to go beyond our proposal and extimate the amount of 
lithium in each sample.
The results obtained were in accordance with the values obtained from the Danish Institute that provided the samples. MEIO INVENTADO

@@
